An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates

### 2022-06-27 poly-embed-journal/poly-embed-journal.tex
Post meeting brain mush

  To see $\H_\Y \subseteq \H^*$, if there was a $y \in \Y$ such that $H^+_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, it is either because it is redundant or $H^+_y \not \subseteq \hyp(g_{L(\R)})$.
  In the first case, there must be some $\epsilon > 0$ such that $\{(x,c) \mid x_y \geq \epsilon\} \in \H^*$, which cannot happen as $g_{L(\R)}$ is concave and finite (namely, nonnegative) at $x = e_y$.
  In the second case, we have a point $x \not \in \reals^\Y_+$ with $g_{L(\R)}(x) > -\infty$.
  This contradicts the fact that $g_{L(\R)}$ is only finite on $\reals^\Y_+$ by construction with the convex indicator. \raft{TODO: Couldn't follow this last line, since $\delta$ can be $\infty$\jessie{does it help that it's minus the convex indicator?} 6/20/2022: still don't get it, sorry :-)... \jessie{Yeah I changed the sentence entirely.  better? \# burn}}. 
\end{proof}

@@

\begin{claim}\label{claim:pi-preserves-dim}
	Consider $L, \V^*$ satisfying Assumption~\ref{assum:V-star-exists-infinite}. %, let $F_v = H_v \cap \hyp(g_\V)$.
%	For all $v \in L(\R)$, define $F_v$ as the face of $g_{L(\R)}=g_{\V^*}$ generated by $H_v$.
	Then $\dim(F_v) = \dim(\pi(F_v))$.
\end{claim}
\begin{proof}
	\jessiet{Can use a careful check}
  \raft{FYI, some of the steps were out of order I think}
	\btw{See \url{http://homepages.cae.wisc.edu/~linderot/classes/ie418/lecture11.pdf} for def of AI}
	%\jessie{just need to argue that you never needed $c$ (i.e., $c$ was always just $g(x)$).  Maybe: take an affine basis, project, and show it's an affine basis.}
	Recall from Definition~\ref{def:face} that the dimension of a polytope to be the dimension of its affine hull.
	Suppose we are given $|\Y|+1$ affinely independent vectors $z_i$ in $F_v$. 
	We will show that their projections $\{\pi(z_i)\}_i$ are affinely independent, giving the result.

@@

\begin{claim}\label{claim:f-min-affine}
	%Let $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$ be a polyhedral concave function with $\dom(f_\V) = \simplex$.
	For any polyhedral concave function with domain on $\simplex$, there is some finite set $\V \subseteq \reals^\Y_+$ such that the function can be written $f_\V(p) = \min_{v\in\V} \inprod{p}{v} - \delta(p \mid \simplex)$.  %\raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
  We will think of $f_\V$ as defined $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(f_\V) = \simplex$.
  For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
  If $p \not \in \simplex$, then $f_\V(p) = -\infty$ and inner products are not used to compute $f_\V$.
  Moreover, since $f_\V$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
	\begin{align*}
	f_\V(p) %&= h(p) - \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

\noindent
This allows us to project $g_\V$ from $\reals^\Y_+$ to the $\simplex$.
\begin{lemma}\label{cor:f-matches-g-on-simplex}
  Consider a finite set $\V \subset \reals^\Y_+$.
  For all $p \in \simplex$, we have $f_\V(p) = g_\V(p)$.
\end{lemma}
%Given the function $f_\V$, we consider $g_\V$ to be its extension and $L$ such that $\risk L_+ = g_\V$ and $\risk L = f_\V$.
Now, we define the function $\theta(v) = \{p \in \simplex \mid \inprod{v}{p} = f_\V(p)\}$ as the level sets of the loss vector $v \in \V$. %, and the set $\Theta_{\V^*} := \{\theta(v) \mid v \in \V^*\}$ to be the set of level sets uniquely defined by a given set of loss vectors.

@@

  Fix $p \in \simplex$.
  \begin{align*}
    p \in \theta(v)
    &\iff \inprod{v}{p} = f_{\V^*}(p) & \text{Definition of $\theta$}\\
    &\iff \inprod{v}{p} = \min_{v' \in {\V^*}}\inprod{v'}{p} & \text{$f_{\V^*}=g_{\V^*}$ on $\simplex$ (Cor.~\ref{cor:f-matches-g-on-simplex})}\\
    &\iff v \in \argmin_{v' \in {\V^*}}\inprod{v'}{p} &\\
    &\iff p \in \pi(F_v)~. & \text{Lemma~\ref{lem:projected-faces-iff-support-iff-argmin}}
  \end{align*}
\end{proof}

@@

\lemmaX*
\begin{proof}
	%\jessiet{Get to Assumption~\ref{assum:V-star-exists-infinite} quickly}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
Since $\risk L$ is polyhedral and concave there exists a finite set $\V$ such that $\risk L = f_\V$ by Claim~\ref{claim:f-min-affine}.
\jessie{I think what follows skips some details}
Moreover as $\V$ is finite, $g_{\V} = \risk L_+$ (by $1$-homogeneity following from the form of $g_{\V}$) is also polyhedral and concave, and $f_\V = g_\V$ on $\simplex$.
Therefore, there is a $\V^* \subseteq \V$ such that $L$ and $\V^*$ satisfy Assumption~\ref{assum:V-star-exists-infinite} by Proposition~\ref{prop:poly-risk-assumption-satisfied}.
Moreover, any set $\R$ such that $\V^* = L(\R)$ is a finite representative set for $L$, and such a set must exist by the existence and construction of $\V^*$.
%Given the form of $\risk L$ written as a pointwise max of finitely many affine functions (which are loss vectors) in $\V$, the loss $L$ has a finite representative set. 
%Moreover, by Claim~\ref{claim:gV-and-fV}, $g_\V = f_\V$ on the simplex.

% for a finite set $\V$ by Claim~\ref{claim:f-min-affine}.
%There is a polyhedral concave function $g_\V$ on $\reals^\Y_+$ matching $f_\V$ on $\simplex$ by Corollary~\ref{cor:f-matches-g-on-simplex}.
%Moreover, consider $g_\V = \risk L_+$\jessiet{is it clear enough that $g_\V = \risk L_+$?}, and observe that $\risk L_+$ matches $\risk L$ on $\simplex$ as well.
%By Corollary~\ref{cor:unique-set-loss-vectors-defining-facets}, we then have a finite set $\V^*$ of smallest cardinality such that $f_\V = f_{\V^*}$ and $g_\V = g_{\V^*}$.
%Consider $\R^* \subseteq \R$ such that $\V^* = L(\R^*)$.
%First, observe that $\R^*$ is representative for $L$ as a corollary of Claim~\ref{claim:projected-faces-cover-RY-iff-representative}.
%Moreover, consider the following set of level sets of $f_{\V^*}$, $\Theta_{\V^*} = \{\theta(v) \mid v \in \V^*\}$.

>>>>

  \jessiet{1. Show it's a facet. 2a. Show $H_y$ is a face. 2b. }
  To show that $\H_\Y \subseteq \H^*$, it suffices from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) to show that each $F_y := H_y \cap \hyp(g_{L(\R)})$ is a facet.
  First, observe that $F_y$ is a face of $\hyp(g_{L(\R)})$ as $H_y$ supports $\hyp(g_{L(\R)})$ by construction of $g_{L(\R)}$ with the convex indicator. \jessie{too much hand-waving?}
  Moreover, $F_y = H_{(e_y, 0)} \cap \hyp(g_{L(\R)}) = \{(x,c) \mid x_y \geq 0\} \cap \hyp(g_{L(\R)})$ is a $|\Y|$-dimensional face as it contains $\{(x,c) \mid x_y \geq 0, c \leq 0\}$, which is $|\Y|$-dimensional.
%  To see $\H_\Y \subseteq \H^*$, if there was a $y \in \Y$ such that $H^+_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, it is either because it is redundant or $H^+_y \not \subseteq \hyp(g_{L(\R)})$.
%  In the first case, there must be some $\epsilon > 0$ such that $\{(x,c) \mid x_y \geq \epsilon\} \in \H^*$, which cannot happen as $g_{L(\R)}$ is concave and finite (namely, nonnegative) at $x = e_y$.
%  In the second case, we have a point $x \not \in \reals^\Y_+$ with $g_{L(\R)}(x) > -\infty$.
%  This contradicts the fact that $g_{L(\R)}$ is only finite on $\reals^\Y_+$ by construction with the convex indicator. \raft{TODO: Couldn't follow this last line, since $\delta$ can be $\infty$\jessie{does it help that it's minus the convex indicator?} 6/20/2022: still don't get it, sorry :-)... \jessie{Yeah I changed the sentence entirely.  better? \# burn}}. 
\end{proof}

@@

\begin{claim}\label{claim:pi-preserves-dim}
	Consider $L, \V^*$ satisfying Assumption~\ref{assum:V-star-exists-infinite}. %, let $F_v = H_v \cap \hyp(g_\V)$.
%	For all $v \in L(\R)$, define $F_v$ as the face of $g_{L(\R)}=g_{\V^*}$ generated by $H_v$.
	Then $\dim(F_v) = \dim(\pi(F_v))$.
\end{claim}
\begin{proof}
	\btw{See \url{http://homepages.cae.wisc.edu/~linderot/classes/ie418/lecture11.pdf} for def of AI}
	%\jessie{just need to argue that you never needed $c$ (i.e., $c$ was always just $g(x)$).  Maybe: take an affine basis, project, and show it's an affine basis.}
	Recall from Definition~\ref{def:face} that the dimension of a polytope to be the dimension of its affine hull.
	Suppose we are given $|\Y|+1$ affinely independent vectors $z_i$ in $F_v$. 
	We will show that their projections $\{\pi(z_i)\}_i$ are affinely independent, giving the result.

@@

\begin{definition}\label{def:f-polyhedral}
	%Let $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$ be a polyhedral concave function with $\dom(f_\V) = \simplex$.
	For any polyhedral concave function $f_\V$ with domain on $\simplex$, we denote the function $f_\V(p) = \min_{v\in\V} \inprod{p}{v} - \delta(p \mid \simplex)$, where $\V \subset \reals^\Y_+$ is a finite set. 
\end{definition}

\noindent
We now make a few observations about the extensions of $1$-homogeneous polyhedral functions.

\begin{claim}\label{claim:extended-poly-risk-poly}
	For any minimizable function $L$ such that $\risk L$ is polyhedral, its extension $\risk L_+$ is also polyhedral.
\end{claim}
\begin{proof}
  We will think of $\risk L$ as defined $\risk L:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(\risk L) = \simplex$.
For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
If $p \not \in \simplex$, then $\risk L(p) = -\infty$ and inner products are not used to compute $\risk L$.
Moreover, since $\risk L$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
\begin{align*}
\risk L(p) 
&= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
&= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
\end{align*}
Therefore, it is positively $1$-homogeneous, and its extension $\risk L_+$ is polyhedral immediately.
\end{proof}

%\noindent
%The construction of $f_\V$ allows us to project $g_\V$ from $\reals^\Y_+$ to the $\simplex$.
%\begin{lemma}\label{cor:f-matches-g-on-simplex}
%  Consider a finite set $\V \subset \reals^\Y_+$.
%  For all $p \in \simplex$, we have $f_\V(p) = g_\V(p)$.
%\end{lemma}
%%Given the function $f_\V$, we consider $g_\V$ to be its extension and $L$ such that $\risk L_+ = g_\V$ and $\risk L = f_\V$.
Now, we define the function $\theta(v) = \{p \in \simplex \mid \inprod{v}{p} = f_\V(p)\}$ as the level sets of the loss vector $v \in \V$. %, and the set $\Theta_{\V^*} := \{\theta(v) \mid v \in \V^*\}$ to be the set of level sets uniquely defined by a given set of loss vectors.

@@

  Fix $p \in \simplex$.
  \begin{align*}
    p \in \theta(v)
    &\iff \inprod{v}{p} = f_{\V^*}(p) & \text{Definition of $\theta$}\\
    &\iff \inprod{v}{p} = \min_{v' \in {\V^*}}\inprod{v'}{p} & \text{$f_{\V^*}=g_{\V^*}$ on $\simplex$ (Claim~\ref{claim:gV-and-fV})}\\
    &\iff v \in \argmin_{v' \in {\V^*}}\inprod{v'}{p} &\\
    &\iff p \in \pi(F_v)~. & \text{Lemma~\ref{lem:projected-faces-iff-support-iff-argmin}}
  \end{align*}
\end{proof}

@@

\lemmaX*
\begin{proof}
	%\jessiet{Get to Assumption~\ref{assum:V-star-exists-infinite} quickly}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
%\jessie{$\risk L$ to $\risk L_+$.  Change Claim 6: if you have a minimizable loss and $\risk L$ is polyhedral, then $\risk L_+$ is polyhedral.}
Since $\risk L$ is polyhedral, so is $\risk L_+$ by Claim~\ref{claim:extended-poly-risk-poly}.
Therefore, we have satisfied the requirements of Proposition~\ref{prop:poly-risk-assumption-satisfied}, and can conclude there is a $\V$ such that $\risk L_+ = g_{\V}$ satisfying Assumption~\ref{assum:V-star-exists-infinite}.
Moreover, there must be a finite set $\R$ such that $\V = L(\R)$ by construction, which is representative as some element of $\R$ generates the Bayes risk for every $p \in \simplex$ as $\risk L = \min_{v \in L(\R)}\inprod{p}{v} - \delta\{p \mid \simplex\}$.
%Since $\risk L$ is polyhedral and concave there exists a finite set $\V$ such that $\risk L = f_\V$ by Claim~\ref{claim:f-min-affine}.
%Moreover as $\V$ is finite, $g_{\V} = \risk L_+$ (by $1$-homogeneity following from the form of $g_{\V}$) is also polyhedral and concave, and $f_\V = g_\V$ on $\simplex$.
%Therefore, there is a $\V^* \subseteq \V$ such that $L$ and $\V^*$ satisfy Assumption~\ref{assum:V-star-exists-infinite} by Proposition~\ref{prop:poly-risk-assumption-satisfied}.
%Moreover, any set $\R$ such that $\V^* = L(\R)$ is a finite representative set for $L$, and such a set must exist by the existence and construction of $\V^*$.
%%Given the form of $\risk L$ written as a pointwise max of finitely many affine functions (which are loss vectors) in $\V$, the loss $L$ has a finite representative set. 
%%Moreover, by Claim~\ref{claim:gV-and-fV}, $g_\V = f_\V$ on the simplex.
